Generalized Plackett-Luce Likelihoods

### 2022-09-27 inst/jss_hyper3.Rnw
issue #161 issue #162

<<>>=
LL <- hyper3()
LL[c(a=1)] <- 1
LL[c(a=3, b=2, c=1)] <- -1
LL
@

@@

<<>>=
a <- 1/2 # null
(S_delta <- log( a*(1-a)/(1+a))- log(3-2*sqrt(2)))
@

@@

<<>>=
pchisq(-2*S_delta,df=1,lower.tail=FALSE)
@

@@

\begin{figure}[htbp]
  \begin{center}
<<figaba>>=
a <- seq(from=0,by=0.005,to=1)
S <- function(a){log(a*(1-a)/((1+a)*(3-2*sqrt(2))))}
plot(a,S(a),type='b')
abline(h=c(0,-2))
abline(v=c(0.02438102,0.9524271),col='red')
abline(v=sqrt(2)-1)
@
\caption{A\label{ABA_likelihood} support function for $p_a$ with observation $a\succ b\succ a$}
\end{center}
\end{figure}

@@

<<label=converttosupp3,cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table, decreasing=TRUE, give.supp=FALSE)
javelin_vector
@

@@

\begin{figure}[htbp]
  \begin{center}
<<testthejav,cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj, pch=16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M), type="b")
abline(h=c(0, -2))
abline(v=0, lty=2)
abline(v=log(0.32062833/0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected}
\end{center}
\end{figure}

@@

<<showconstructortable>>=
constructor_2021_table[,1:9]
@

@@

<<definecombinedlikelihoodfunction,cache=TRUE>>=
H <- psubs(constructor_2020, "Merc", "Merc2020") + psubs(constructor_2021, "Merc", "Merc2021")
@

>>>>

<<>>=
LL <- hyper3()
LL[c(a = 1)] <- 1
LL[c(a = 3, b = 2, c = 1)] <- -1
LL
@

@@

<<>>=
a <- 1/2 # null
(S_delta <- log(a * (1 - a)/(1 + a)) - log(3 - 2 * sqrt(2)))
@

@@

<<>>=
pchisq(-2 * S_delta, df=1, lower.tail=FALSE)
@

@@

\begin{figure}[htbp]
  \begin{center}
<<figaba>>=
a <- seq(from=0, by=0.005, to=1)
S <- function(a){log(a * (1 - a) / ((1 + a) * (3 - 2 * sqrt(2))))}
plot(a, S(a), type='b')
abline(h=c(0, -2))
abline(v=c(0.02438102, 0.9524271), col='red')
abline(v=sqrt(2) - 1)
@
\caption{A\label{ABA_likelihood} support function for $p_a$ with observation $a\succ b\succ a$}
\end{center}
\end{figure}

@@

<<label=converttosupp3, cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table, decreasing=TRUE, give.supp=FALSE)
javelin_vector
@

@@

\begin{figure}[htbp]
  \begin{center}
<<testthejav, cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj, pch=16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M), type="b")
abline(h=c(0, -2))
abline(v=0, lty=2)
abline(v=log(0.32062833 / 0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected}
\end{center}
\end{figure}

@@

<<showconstructortable>>=
constructor_2021_table[, 1:9]
@

@@

<<definecombinedlikelihoodfunction,cache=TRUE>>=
H <- (
      psubs(constructor_2020, "Merc", "Merc2020") +
      psubs(constructor_2021, "Merc", "Merc2021")
     )
@
